---
title: "Unusual Data Vizzes"
author: "Zack Wixom"
output: html_document
---

### Word of Warning

These are unusual data visualizations, you should in general stick to the Power 6 charts (bar, line, scatter, table, text, and pie) but these are just super cool and fun ones.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Density Charts

First group of data vizzes are Density Charts. 

###Averages are dangerous

```{r}

library(tidyverse)

houses <- read_csv(here::here("data", "cleaned_parcels.csv"))

houses <- houses %>%
  select(parcel_number, property_add, market_value, year_built) %>%
  separate(property_add, c("address", "city"), sep = ",") %>%
  mutate(city = trimws(city)) %>%
  filter(market_value > 0,
         year_built > 0,
         city %in% c("WOODLAND", "STRAWBERRY", "MIDWAY", "KAMAS", "HEBER"))


```

### Histograms
```{r}
ggplot(houses) +
  geom_histogram(aes(x = market_value)) + 
  xlim(0, 1000000)

```

## Density

```{r}
ggplot(houses) +
  geom_density(aes(x = market_value), fill = "grey50") + 
  xlim(0, 1000000)

```

## Multiple Densities
```{r}
ggplot(houses) +
  geom_density(aes(x = market_value, fill = city), alpha = 0.3) + 
  xlim(0, 1000000)

```

## Multiple with Facets
```{r}
ggplot(houses) +
  geom_density(aes(x = market_value), fill = "grey50") + 
  facet_grid(~city) +
  xlim(0, 1000000)

```

## Ridgelines
```{r}

# install.packages("ggridges")

library(ggridges)

ggplot(houses) +
  geom_density_ridges(aes(y = city, x = market_value)) + 
  xlim(0, 1000000)


```

## Violine

```{r}
ggplot(houses) +
  geom_violin(aes(x = city, y = market_value)) +
  ylim(0, 1000000) + 
  coord_flip()


```

## Box Plot

```{r}
ggplot(houses) +
  geom_boxplot(aes(x = city, y = market_value)) +
  ylim(0, 1000000) + 
  coord_flip()

```


## Jitter
```{r}
ggplot(houses) +
  geom_jitter(aes(x = city, y = market_value)) +
  ylim(0, 1000000) + 
  coord_flip()


```

## Frankenstein

```{r}
ggplot(houses, aes(x = city, y = market_value)) +
  geom_boxplot(alpha = .5, color = "blue") +
  geom_violin(alpha = .5, color = "red") +
  geom_jitter(alpha = 0.1) +
  ylim(0, 1000000) + 
  coord_flip()

```


## Combination with charts in Margins

Source: https://www.r-graph-gallery.com/277-marginal-histogram-for-ggplot2.html

## Rug Plot

Wasn't able to do


## Heat Maps

```{r}
houses %>%
  filter(city %in% c("MIDWAY", "HEBER"),
         year_built >= 2010,
         year_built <= 2018) %>%
  group_by(city, year_built) %>%
  summarise(med_price = median(market_value)) %>%
  mutate(med_price = med_price/1000) %>%
  ggplot(aes(y = year_built, x = city, fill = med_price)) + 
  geom_tile() +
  geom_text(aes(label = scales::label_dollar(suffix = "K")(med_price)))

```

## Scatter not helpful with this data

```{r}
ggplot(houses, aes(x = year_built, y = market_value)) +
  geom_point() + 
  ylim(0, 1000000) + 
  xlim(1975, 2018)

```

## Add some Jitter to it

```{r}
ggplot(houses, aes(x = year_built, y = market_value)) +
  geom_jitter(alpha = .5) + 
  ylim(0, 1000000) + 
  xlim(1975, 2018)


```

## Hexbin

```{r}
ggplot(houses, aes(x = year_built, y = market_value)) +
  geom_hex() + 
  ylim(0, 1000000) + 
  xlim(1975, 2018)

```

## Contour Map

```{r}
ggplot(houses, aes(x = year_built, y = market_value)) +
  geom_density2d() + 
  ylim(0, 1000000) + 
  xlim(1975, 2018)


```

## Density 2D

```{r}
ggplot(houses, aes(x = year_built, y = market_value)) +
  stat_density_2d(aes(fill = ..level..), geom = "polygon") +
  ylim(0, 1000000) + 
  xlim(1975, 2018)


```

Customizing Options: https://www.r-graph-gallery.com/2d-density-plot-with-ggplot2.html#distr


# Eye Candy

Super cool ones. require additional packages to do

## Sankey

Can be useful and show flows. Flowing from one thing to another.

## Chord

Sankey chart with a polar axis. hard to pin point the data story

## Radar or Spider Charts

Common in politics and Sociology. Lots of criticism around this chart type.

## Word Clouds

Clients love these, not amazingly useful but cool. Ideal to keep them all horizontal. Also use familiar sizes and color/saturation for the words.

## Network graphs

sometimes called a hairball. Can be too crazy to be useful. Bad for presenting data. Mediocre for analysis.

## Dendrogram

Common with decision tree. Lots of cool R packages that make this loop nice and work with decision trees in R. 


# Charts by Other Names

## Streamgraph or steamgraph

```{r}
# First staked area
houses %>%
  count(city, year_built) %>%
  filter(between(year_built, 1975, 2018)) %>%
  ggplot(aes(x = year_built, y = n, fill = city)) + 
  geom_area()

```

```{r}
devtools::install_github("hrbrmstr/streamgraph")
library(streamgraph)
houses %>%
  count(city, year_built) %>%
  filter(between(year_built, 1975, 2018)) %>%
  streamgraph(key="city", value="n", date="year_built")


```


## Lollipop

Nice break in bar charts

```{r}
houses %>%
  filter(year_built == 2017) %>%
  count(city) %>%
  arrange(n) %>%
  mutate(city = fct_inorder(city)) %>%
  ggplot() + 
  geom_bar(aes(y = city, x = n), stat = "identity")

ggplot(houses) + 
  geom_bar(aes(y=city))

```
```{r}
houses %>%
  filter(year_built == 2017) %>%
  count(city) %>%
  arrange(n) %>%
  mutate(city = fct_inorder(city)) %>%
  ggplot() + 
  geom_segment(aes(x = 0, xend = n, y = city , yend = city), color = "grey50", size = 2) +
  geom_point(aes(x = n, y = city), color = "red", size = 8) 


```

