---
title: "ggplot2 - inclass work"
author: "Zack Wixom"
output: html_notebook
---

## ggplot2
written by Hadley Wickam. Wrote it as a grad project and got the name from the term _grammar of graphics_.

## Key Takeaways of ggplot

- Works best with "tidy data"
- You specify the mapping (encoding) of the data to different aesthetics (channels) and plot the geometric objects/geoms (marks = _ink on screen_)


```{r load packages}

library(tidyverse)
library(ggplot2) # just the ggplot packages

```



ggplot
useful stuff:
- __alpha__ = density or transparency


```{r data}

# read in Data
diamonds <- diamonds

# Clean Data
diamonds

```

## ggplot Function

``` {r ggplot function}

# Analysis & Visualization

# ggplot function
ggplot(data = diamonds, mapping = aes()) +
  geom_point(mapping = aes(x = carat, y = price)) +
  geom_vline(mapping = aes(xintercept = 3)) +
  geom_hline(mapping = aes(yintercept = 5000))
  
ggplot(diamonds) +
  geom_point(aes(carat, price, size = depth *.5), color = "#32a87b", alpha = .2) +
  geom_vline(aes(xintercept = 3), color = "#a87d32") +
  geom_hline(aes(yintercept = 5000), color = "#a87d32") +
  scale_size_continuous(range = c(.5, 1)) +
  scale_y_continuous(labels = dollar_format(suffix = "K", scale = .001), breaks = c(5000, 10000, 15000)) +
  theme_wsj() +
  theme(
    legend.position = "none",
    axis.title = element_blank(),
    text = element_text("Avenir", color = "grey60"),
    title = element_text("Avenir", color = "grey60")
  ) +
  labs(
    title = "Price of Diamonds",
    subtitle = "Per Carat",
    caption = "Source: ggplot"
  )
  

```

Also can add a coordinate plane:
- coord_flip
- coord_map
- coord_polar

Also a theme
- classic is default

Also a Facet
- facet_wrap = only one facet variable
- facet_grid = can do more than one variable to facet by
```{r}

diamonds %>% 
  mutate(
    trans = ifelse(color == "D", .8, 0.2)
  ) %>% 
  ggplot() +
  geom_point(aes(x = carat, y = price, color = color, size = depth, alpha = trans)) +
  coord_polar() +
  theme_minimal() +
  facet_grid(~color)
  
```
Now for the Homework I need to recreate one of my vizzes using ggplot

```{r}
# load packages for excel data loading and here for data path
library(readxl)
library(here)

sales <- read_xlsx(here("data_r", "sales_rep_data.xlsx"))

sales <- sales %>%
  rename(sales_rep = `Sales Rep`)

average <- data.frame(list("Average Total Sales", 37480.26, 37480.26))


sales_ave <- sales %>% 
  group_by(sales_rep) %>% 
  summarize(
    ave_sales = mean(Sales)
  ) %>%
  mutate(
    total_ave_sales = ifelse(sales_rep != "Ellie-Rose Wharton",mean(sort(ave_sales)[-c(1, length(ave_sales))]), ave_sales)
  )

names(average) <- names(sales_ave)

sales_simp <- sales_ave %>% 
  rbind(average) %>%
  arrange(desc(total_ave_sales)) %>% 
  head(3)

sales_simp[3,1] <- "All Other Reps"

library(scales)

my_colors <- c("#306165", "#A78D4E", "#A24C52")

my_white <- "#F8F8F8"

h <- max(sales_simp[2,3])

sales_simp %>%
  group_by(total_ave_sales) %>% 
  ggplot() +
  geom_col(aes(x = reorder(sales_rep, -total_ave_sales), y = total_ave_sales, fill = sales_rep), show.legend = FALSE) +
  scale_fill_manual("sales_rep", values = my_colors) +
  scale_y_continuous(label = label_number(prefix = "$", suffix = "K", scale = 1/1e3), breaks = c(20000, 40000, 60000,80000)) +
  labs(
    title = "Wharton Outperforms All Other Reps",
    subtitle = "Average sales over last two years",
    y = element_blank(),
    x = element_blank()
  ) +
  theme_few() +
  theme(
    plot.background = element_rect(fill = "#F8F8F8"),
    panel.background = element_blank(),
    panel.border = element_blank(),
    text = element_text(family = "Avenir", size = 16),
    axis.ticks = element_blank(),
    panel.spacing.y = unit(c(0, 0, 0, 0), "null")
  ) +
  geom_hline(aes(yintercept = h), color = "#A78D4E")
  



```


gapminder exercise

```{r}

library(gapminder)
gapminder <- gapminder


gapminder %>% 
  filter(year == 1997) %>% 
  ggplot(aes(gdpPercap, lifeExp, size = pop)) +
  geom_point(aes(color = continent)) +
  scale_x_log10()


```

